---
title: "Results"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(flextable)
library(dplyr)

# Source all data prep and model results
source("analyses.R")

### Formatting helpers ###

# β coefficient 
fmt_b <- function(x) formatC(as.numeric(x$coef), format = "f", digits = 4)

# 95% CI as "lo to hi" for inline text
fmt_ci_text <- function(x) {
  ci <- as.numeric(x$ci)
  paste0(formatC(ci[1], format = "f", digits = 4), " to ",
         formatC(ci[2], format = "f", digits = 4))
}

# 95% CI as "lo-hi" for table cells
fmt_ci_dash <- function(x) {
  ci <- as.numeric(x$ci)
  paste0(formatC(ci[1], format = "f", digits = 4), "-",
         formatC(ci[2], format = "f", digits = 4))
}

# "Coef (lo-hi)" combined cell for tables
fmt_coef_ci <- function(x) paste0(fmt_b(x), " (", fmt_ci_dash(x), ")")

# p-value for inline text: returns "=0.0450" or "<0.0001"
fmt_p <- function(p) {
  p <- as.numeric(p)[1]
  if (is.na(p))     return("=NA")
  if (p < 0.0001)   return("<0.0001")
  paste0("=", formatC(p, format = "f", digits = 4))
}

# p-value for table cells, with significance stars
fmt_p_star <- function(p) {
  p <- as.numeric(p)[1]
  if (is.na(p)) return("")
  base <- formatC(p, format = "f", digits = 4)
  if      (p < 0.001) paste0(base, "***")
  else if (p < 0.01)  paste0(base, "**")
  else if (p < 0.05)  paste0(base, "*")
  else                base
}

# Look up FDR p-value (formatted) from a results tibble by model name
get_fdr <- function(fdr_df, model_name) {
  p <- fdr_df$p_fdr[fdr_df$model == model_name]
  if (length(p) == 0) return("=NA")
  fmt_p(p)
}

# Look up FDR p-value (raw numeric) — for table formatting
get_fdr_raw <- function(fdr_df, model_name) {
  p <- fdr_df$p_fdr[fdr_df$model == model_name]
  if (length(p) == 0) return(NA_real_)
  as.numeric(p)[1]
}

# Marginal / conditional R² from LME models
fmt_r2m <- function(x) formatC(as.numeric(x$marginal_r2),    format = "f", digits = 4)
fmt_r2c <- function(x) formatC(as.numeric(x$conditional_r2), format = "f", digits = 4)

# R² from LM models
fmt_r2 <- function(x) formatC(as.numeric(x$r2), format = "f", digits = 4)

# Mean ± SD 
fmt_ms <- function(x) {
  paste0(formatC(mean(x, na.rm = TRUE), format = "f", digits = 2),
         " \u00b1 ",
         formatC(sd(x,   na.rm = TRUE), format = "f", digits = 2))
}

### Derived demographic stats ###

n_total <- nrow(summary_df)
n_sa    <- sum(summary_df$group == "Superager",     na.rm = TRUE)
n_nsa   <- sum(summary_df$group == "Non-superager", na.rm = TRUE)
pct_sa  <- round(n_sa / n_total * 100, 1)

is_female   <- function(x) tolower(as.character(x)) == "female"
n_f_total   <- sum(is_female(summary_df$sex),                                              na.rm = TRUE)
n_f_sa      <- sum(is_female(summary_df$sex[summary_df$group == "Superager"]),             na.rm = TRUE)
n_f_nsa     <- sum(is_female(summary_df$sex[summary_df$group == "Non-superager"]),         na.rm = TRUE)
pct_f_total <- round(n_f_total / n_total * 100, 1)
pct_f_sa    <- round(n_f_sa    / n_sa    * 100, 1)
pct_f_nsa   <- round(n_f_nsa   / n_nsa   * 100, 1)

mean_age <- formatC(mean(summary_df$age_1, na.rm = TRUE), format = "f", digits = 2)
sd_age   <- formatC(sd(summary_df$age_1,   na.rm = TRUE), format = "f", digits = 2)

# Minimum p across RAVLT slope models (for the "ps >" summary statement)
min_ravlt_slope_p <- min(
  sapply(ravlt_slope_model_stats, function(x) as.numeric(x$p_value)[1]),
  na.rm = TRUE
)
# Floor to 1 d.p. so the "ps >" statement is conservative
min_ravlt_slope_p_floor <- formatC(floor(min_ravlt_slope_p * 10) / 10,
                                   format = "f", digits = 1)
```

Our overall cohort included `r n_total` participants with `r pct_f_total`% female and a mean baseline age of `r mean_age` (SD `r sd_age`) years. `r pct_sa`% of the cohort met longitudinal superager criteria. Superagers and non-superagers did not vary on baseline age (p`r fmt_p(age_baseline_p)`), years of education (p`r fmt_p(yoe_p)`), baseline MMSE scores (p`r fmt_p(mmse_p)`), or follow-up duration (p`r fmt_p(followup_duration_p)`); however, superagers were more likely to be female (p`r fmt_p(sex_p)`). These demographic statistics are summarized in Table 2.

**Table 2** Demographic characteristics

```{r table2}
tbl2 <- tibble(
  Variable = c(
    "Number of participants",
    "Number females (percentage)",
    "Age baseline \u00b1 SD",
    "Follow-up duration (years) \u00b1 SD",
    "Education (years) \u00b1 SD",
    "MMSE baseline\u00b1 SD"
  ),
  `Whole cohort` = c(
    as.character(n_total),
    paste0(n_f_total, " (", pct_f_total, "%)"),
    fmt_ms(summary_df$age_1),
    fmt_ms(summary_df$followup_duration),
    fmt_ms(summary_df$YoE),
    fmt_ms(summary_df$mmse_1)
  ),
  `Non-superagers` = c(
    as.character(n_nsa),
    paste0(n_f_nsa, " (", pct_f_nsa, "%)"),
    fmt_ms(summary_df$age_1[summary_df$group         == "Non-superager"]),
    fmt_ms(summary_df$followup_duration[summary_df$group == "Non-superager"]),
    fmt_ms(summary_df$YoE[summary_df$group           == "Non-superager"]),
    fmt_ms(summary_df$mmse_1[summary_df$group        == "Non-superager"])
  ),
  `Superagers` = c(
    as.character(n_sa),
    paste0(n_f_sa, " (", pct_f_sa, "%)"),
    fmt_ms(summary_df$age_1[summary_df$group         == "Superager"]),
    fmt_ms(summary_df$followup_duration[summary_df$group == "Superager"]),
    fmt_ms(summary_df$YoE[summary_df$group           == "Superager"]),
    fmt_ms(summary_df$mmse_1[summary_df$group        == "Superager"])
  ),
  `p-value` = c(
    "--",
    fmt_p_star(sex_p),
    fmt_p_star(age_baseline_p),
    fmt_p_star(followup_duration_p),
    fmt_p_star(yoe_p),
    fmt_p_star(mmse_p)
  )
)

flextable(tbl2) |>
  bold(part = "header") |>
  autofit()
```

*Abbreviations: SD standard deviation, MMSE Mini Mental State Examination*

**Structure-function coupling, superager status, and episodic memory:** After running the elastic net models, we longitudinally examined whether SFC varied by superager status. We found that superagers, compared to non-superagers, consistently showed nominally higher SFC globally (β=`r fmt_b(sfc_superager_weighted)`, 95% CI `r fmt_ci_text(sfc_superager_weighted)`, p`r fmt_p(sfc_superager_weighted$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_weighted_mean")`, marginal r²=`r fmt_r2m(sfc_superager_weighted)`, conditional r²=`r fmt_r2c(sfc_superager_weighted)`), in heteromodal regions (β=`r fmt_b(sfc_superager_hmod)`, 95% CI `r fmt_ci_text(sfc_superager_hmod)`, p`r fmt_p(sfc_superager_hmod$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_hmod")`, marginal r²=`r fmt_r2m(sfc_superager_hmod)`, conditional r²=`r fmt_r2c(sfc_superager_hmod)`), and for regions important for memory including the DMN (β=`r fmt_b(sfc_superager_dmn)`, 95% CI `r fmt_ci_text(sfc_superager_dmn)`, p`r fmt_p(sfc_superager_dmn$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_dmn")`, marginal r²=`r fmt_r2m(sfc_superager_dmn)`, conditional r²=`r fmt_r2c(sfc_superager_dmn)`) and ECN (β=`r fmt_b(sfc_superager_control)`, 95% CI `r fmt_ci_text(sfc_superager_control)`, p`r fmt_p(sfc_superager_control$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_control")`, marginal r²=`r fmt_r2m(sfc_superager_control)`, conditional r²=`r fmt_r2c(sfc_superager_control)`). We found a trend toward significance in the SN (β=`r fmt_b(sfc_superager_salience)`, 95% CI `r fmt_ci_text(sfc_superager_salience)`, p`r fmt_p(sfc_superager_salience$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_salience")`, marginal r²=`r fmt_r2m(sfc_superager_salience)`, conditional r²=`r fmt_r2c(sfc_superager_salience)`); however, none of these results survived correction for multiple comparisons. Additionally, we found no significant difference in SFC by superager status in sensory regions (p`r fmt_p(sfc_superager_sensory$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_sensory")`). Figure 3 panel A presents the difference in global SFC between superagers and non-superagers averaged across baseline and follow-up, with warmer colors showing regions with higher SFC in superagers and panel B summarizes these results. 

**Figure 3** Structure-function coupling, episodic memory, and superagers status  

Figure 3 (A) shows a cortical surface and subcortical map of the unthresholded mean difference in global structure-function coupling between superagers and non-superagers, averaged across both timepoints. Warm colors indicate higher coupling in superagers; cool colors indicate lower coupling. Panels (B, C) present linear mixed effects model coefficients and confidence intervals for regional structure-function coupling as a function of superager status (B) and episodic memory (C). Filled circles: p < .05 (uncorrected), open circles: non-significant. * pFDR < .05. Abbreviations: FDR false discovery rate, SFC structure-function coupling, DMN default mode network, ECN executive control network, SN salience network, CI confidence interval.

When we examined SFC and episodic memory, we found that higher SFC was associated with better episodic memory in the whole cohort when measured globally (β=`r fmt_b(ravlt_sfc_weighted)`, 95% CI `r fmt_ci_text(ravlt_sfc_weighted)`, p`r fmt_p(ravlt_sfc_weighted$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_weighted_mean")`, marginal r²=`r fmt_r2m(ravlt_sfc_weighted)`, conditional r²=`r fmt_r2c(ravlt_sfc_weighted)`), in heteromodal regions (β=`r fmt_b(ravlt_sfc_hmod)`, 95% CI `r fmt_ci_text(ravlt_sfc_hmod)`, p`r fmt_p(ravlt_sfc_hmod$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_hmod")`, marginal r²=`r fmt_r2m(ravlt_sfc_hmod)`, conditional r²=`r fmt_r2c(ravlt_sfc_hmod)`), and in the DMN (β=`r fmt_b(ravlt_sfc_dmn)`, 95% CI `r fmt_ci_text(ravlt_sfc_dmn)`, p`r fmt_p(ravlt_sfc_dmn$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_dmn")`, marginal r²=`r fmt_r2m(ravlt_sfc_dmn)`, conditional r²=`r fmt_r2c(ravlt_sfc_dmn)`). Though there was a trend, we found no significant association between ECN (p`r fmt_p(ravlt_sfc_control$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_control")`) and SN (p`r fmt_p(ravlt_sfc_salience$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_salience")`) SFC and episodic memory. Moreover, there was no evidence of an association between sensory region SFC and episodic memory (p`r fmt_p(ravlt_sfc_sensory$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_sensory")`). These results are displayed in Figure 3 panel C. 

We then examined SFC annual change and superager status. We did not find any evidence of significant relationships between superager status and global (p`r fmt_p(sfc_superager_weighted_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_weighted_mean_slope")`), heteromodal (p`r fmt_p(sfc_superager_hmod_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_hmod_slope")`), DMN (p`r fmt_p(sfc_superager_dmn_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_dmn_slope")`), ECN (p`r fmt_p(sfc_superager_control_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_control_slope")`), or sensory SFC annual change (p`r fmt_p(sfc_superager_sensory_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_sensory_slope")`). We found limited evidence of an association between SN SFC annual change and superager status, but this did not survive correction for multiple comparisons (p`r fmt_p(sfc_superager_salience_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_salience_slope")`). With SFC annual change and episodic memory, we found no evidence that SFC annual change in any of the investigated regions was associated with episodic memory (ps > `r min_ravlt_slope_p_floor`). These results are summarized in Table 3.

**Table 3** Associations between structure-function coupling, superager status, and episodic memory

```{r table3}
# Helpers to build data rows for LME and LM models
lme_row <- function(region, model, fdr_df, fdr_key) {
  list(
    Region                 = region,
    `Coefficient (95% CI)` = fmt_coef_ci(model),
    `p-value`              = fmt_p_star(as.numeric(model$p_value)[1]),
    pFDR                   = fmt_p_star(get_fdr_raw(fdr_df, fdr_key)),
    `Marginal r²`     = fmt_r2m(model),
    `Conditional r²`  = fmt_r2c(model)
  )
}

lm_row <- function(region, model, fdr_df, fdr_key) {
  list(
    Region                 = region,
    `Coefficient (95% CI)` = fmt_coef_ci(model),
    `p-value`              = fmt_p_star(as.numeric(model$p_value)[1]),
    pFDR                   = fmt_p_star(get_fdr_raw(fdr_df, fdr_key)),
    `Marginal r²`     = fmt_r2(model),
    `Conditional r²`  = ""
  )
}

# Section header row (spans all columns via merge_at below)
hdr_row <- function(label) {
  list(Region = label, `Coefficient (95% CI)` = "",
       `p-value` = "", pFDR = "",
       `Marginal r²` = "", `Conditional r²` = "")
}

# Section type row (model description label, styled differently from hdr_row)
type_row <- function(label) {
  list(Region = label, `Coefficient (95% CI)` = "",
       `p-value` = "", pFDR = "",
       `Marginal r²` = "", `Conditional r²` = "")
}

tbl3 <- bind_rows(
  # ── Superager status (LME) ──────────────────────────────────────────────
  type_row("Repeated measures linear mixed effects analyses"),
  hdr_row("Superager status"),
  lme_row("Global SFC",      sfc_superager_weighted, sfc_fdr_results, "sfc_weighted_mean"),
  lme_row("Heteromodal SFC", sfc_superager_hmod,     sfc_fdr_results, "sfc_hmod"),
  lme_row("DMN SFC",         sfc_superager_dmn,      sfc_fdr_results, "sfc_dmn"),
  lme_row("ECN SFC",         sfc_superager_control,  sfc_fdr_results, "sfc_control"),
  lme_row("SN SFC",          sfc_superager_salience, sfc_fdr_results, "sfc_salience"),
  lme_row("Sensory SFC",     sfc_superager_sensory,  sfc_fdr_results, "sfc_sensory"),

  # ── Superager status — annual change (LM) ───────────────────────────────
  type_row("Linear regression analyses — annual change"),
  hdr_row("Superager status"),
  lm_row("Global SFC annual change",      sfc_superager_weighted_slope, sfc_slope_fdr_results, "sfc_weighted_mean_slope"),
  lm_row("Heteromodal SFC annual change", sfc_superager_hmod_slope,     sfc_slope_fdr_results, "sfc_hmod_slope"),
  lm_row("DMN SFC annual change",         sfc_superager_dmn_slope,      sfc_slope_fdr_results, "sfc_dmn_slope"),
  lm_row("ECN SFC annual change",         sfc_superager_control_slope,  sfc_slope_fdr_results, "sfc_control_slope"),
  lm_row("SN SFC annual change",          sfc_superager_salience_slope, sfc_slope_fdr_results, "sfc_salience_slope"),
  lm_row("Sensory SFC annual change",     sfc_superager_sensory_slope,  sfc_slope_fdr_results, "sfc_sensory_slope"),

  # ── Episodic memory (LME) ───────────────────────────────────────────────
  type_row("Repeated measures linear mixed effects analyses"),
  hdr_row("Episodic memory"),
  lme_row("Global SFC",      ravlt_sfc_weighted, ravlt_fdr_results, "ravlt_sfc_weighted_mean"),
  lme_row("Heteromodal SFC", ravlt_sfc_hmod,     ravlt_fdr_results, "ravlt_sfc_hmod"),
  lme_row("DMN SFC",         ravlt_sfc_dmn,      ravlt_fdr_results, "ravlt_sfc_dmn"),
  lme_row("ECN SFC",         ravlt_sfc_control,  ravlt_fdr_results, "ravlt_sfc_control"),
  lme_row("SN SFC",          ravlt_sfc_salience, ravlt_fdr_results, "ravlt_sfc_salience"),
  lme_row("Sensory SFC",     ravlt_sfc_sensory,  ravlt_fdr_results, "ravlt_sfc_sensory"),

  # ── Episodic memory — annual change (LME) ───────────────────────────────
  type_row("Repeated measures linear mixed effects analyses — annual change"),
  hdr_row("Episodic memory"),
  lme_row("Global SFC annual change",      ravlt_sfc_weighted_slope, ravlt_slope_fdr_results, "ravlt_sfc_weighted_mean_slope"),
  lme_row("Heteromodal SFC annual change", ravlt_sfc_hmod_slope,     ravlt_slope_fdr_results, "ravlt_sfc_hmod_slope"),
  lme_row("DMN SFC annual change",         ravlt_sfc_dmn_slope,      ravlt_slope_fdr_results, "ravlt_sfc_dmn_slope"),
  lme_row("ECN SFC annual change",         ravlt_sfc_control_slope,  ravlt_slope_fdr_results, "ravlt_sfc_control_slope"),
  lme_row("SN SFC annual change",          ravlt_sfc_salience_slope, ravlt_slope_fdr_results, "ravlt_sfc_salience_slope"),
  lme_row("Sensory SFC annual change",     ravlt_sfc_sensory_slope,  ravlt_slope_fdr_results, "ravlt_sfc_sensory_slope")
)

# Identify type rows (model description) and section header rows
type_labels <- c(
  "Repeated measures linear mixed effects analyses",
  "Linear regression analyses — annual change",
  "Repeated measures linear mixed effects analyses — annual change"
)
type_idx <- which(tbl3$Region %in% type_labels)
hdr_idx  <- setdiff(which(tbl3$`Coefficient (95% CI)` == ""), type_idx)

ft <- flextable(tbl3) |>
  bold(part = "header") |>
  bold(i = hdr_idx) |>
  bg(i = hdr_idx, bg = "#F0F0F0") |>
  italic(i = type_idx) |>
  bg(i = type_idx, bg = "#E8EEF4") |>
  autofit()

# Merge both type and header rows across all columns so labels span the full width
for (r in c(type_idx, hdr_idx)) {
  ft <- merge_at(ft, i = r, j = 1:6)
}

ft
```

*Abbreviations: CI confidence interval, FDR False Discovery Rate, SFC structure-function coupling, DMN default mode network, ECN executive control network, SN salience network*
