---
title: "Results"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(flextable)
library(dplyr)

# Source all data prep and model results
source("analyses.R")

### Formatting helpers ###

# β coefficient 
fmt_b <- function(x) formatC(as.numeric(x$coef), format = "f", digits = 4)

# 95% CI as "lo to hi" for inline text
fmt_ci_text <- function(x) {
  ci <- as.numeric(x$ci)
  paste0(formatC(ci[1], format = "f", digits = 4), " to ",
         formatC(ci[2], format = "f", digits = 4))
}

# 95% CI as "lo-hi" for table cells
fmt_ci_dash <- function(x) {
  ci <- as.numeric(x$ci)
  paste0(formatC(ci[1], format = "f", digits = 4), "-",
         formatC(ci[2], format = "f", digits = 4))
}

# "Coef (lo-hi)" combined cell for tables
fmt_coef_ci <- function(x) paste0(fmt_b(x), " (", fmt_ci_dash(x), ")")

# p-value for inline text: returns "=0.0450" or "<0.0001"
fmt_p <- function(p) {
  p <- as.numeric(p)[1]
  if (is.na(p))     return("=NA")
  if (p < 0.0001)   return("<0.0001")
  paste0("=", formatC(p, format = "f", digits = 4))
}

# p-value for table cells, with significance stars
fmt_p_star <- function(p) {
  p <- as.numeric(p)[1]
  if (is.na(p)) return("")
  base <- formatC(p, format = "f", digits = 4)
  if      (p < 0.001) paste0(base, "***")
  else if (p < 0.01)  paste0(base, "**")
  else if (p < 0.05)  paste0(base, "*")
  else                base
}

# Look up FDR p-value (formatted) from a results tibble by model name
get_fdr <- function(fdr_df, model_name) {
  p <- fdr_df$p_fdr[fdr_df$model == model_name]
  if (length(p) == 0) return("=NA")
  fmt_p(p)
}

# Look up FDR p-value (raw numeric) — for table formatting
get_fdr_raw <- function(fdr_df, model_name) {
  p <- fdr_df$p_fdr[fdr_df$model == model_name]
  if (length(p) == 0) return(NA_real_)
  as.numeric(p)[1]
}

# Marginal / conditional R² from LME models
fmt_r2m <- function(x) formatC(as.numeric(x$marginal_r2),    format = "f", digits = 4)
fmt_r2c <- function(x) formatC(as.numeric(x$conditional_r2), format = "f", digits = 4)

# Mean ± SD
fmt_ms <- function(x) {
  paste0(formatC(mean(x, na.rm = TRUE), format = "f", digits = 2),
         " \u00b1 ",
         formatC(sd(x,   na.rm = TRUE), format = "f", digits = 2))
}

### Derived demographic stats ###

n_total <- nrow(summary_df)
n_sa    <- sum(summary_df$group == "Superager",     na.rm = TRUE)
n_nsa   <- sum(summary_df$group == "Non-superager", na.rm = TRUE)
pct_sa  <- round(n_sa / n_total * 100, 1)

is_female   <- function(x) tolower(as.character(x)) == "female"
n_f_total   <- sum(is_female(summary_df$sex),                                              na.rm = TRUE)
n_f_sa      <- sum(is_female(summary_df$sex[summary_df$group == "Superager"]),             na.rm = TRUE)
n_f_nsa     <- sum(is_female(summary_df$sex[summary_df$group == "Non-superager"]),         na.rm = TRUE)
pct_f_total <- round(n_f_total / n_total * 100, 1)
pct_f_sa    <- round(n_f_sa    / n_sa    * 100, 1)
pct_f_nsa   <- round(n_f_nsa   / n_nsa   * 100, 1)

mean_age <- formatC(mean(summary_df$age_1, na.rm = TRUE), format = "f", digits = 2)
sd_age   <- formatC(sd(summary_df$age_1,   na.rm = TRUE), format = "f", digits = 2)

# Min p and pFDR across SFC change models, excluding SN (reported separately)
min_sfc_slope_p <- min(
  sapply(sfc_slope_model_stats[names(sfc_slope_model_stats) != "sfc_salience_slope"],
         function(x) as.numeric(x$p_value)[1]), na.rm = TRUE)
min_sfc_slope_p_fdr <- min(
  sfc_slope_fdr_results$p_fdr[sfc_slope_fdr_results$model != "sfc_salience_slope"], na.rm = TRUE)
min_ravlt_slope_p <- min(
  sapply(ravlt_slope_model_stats[names(ravlt_slope_model_stats) != "ravlt_sfc_salience_slope"],
         function(x) as.numeric(x$p_value)[1]), na.rm = TRUE)
min_ravlt_slope_p_fdr <- min(
  ravlt_slope_fdr_results$p_fdr[ravlt_slope_fdr_results$model != "ravlt_sfc_salience_slope"], na.rm = TRUE)
fmt_round2 <- function(p) formatC(round(p, 2), format = "f", digits = 2)
```

Our overall cohort included `r n_total` participants with `r pct_f_total`% female and a mean baseline age of `r mean_age` (SD `r sd_age`) years. `r pct_sa`% of the cohort met longitudinal superager criteria. Superagers and non-superagers did not vary on baseline age (p`r fmt_p(age_baseline_p)`), years of education (p`r fmt_p(yoe_p)`), baseline MMSE scores (p`r fmt_p(mmse_p)`), or follow-up duration (p`r fmt_p(followup_duration_p)`); however, superagers were more likely to be female (p`r fmt_p(sex_p)`). These demographic statistics are summarized in Table 1.

**Table 1** Demographic characteristics

```{r table2}
tbl2 <- tibble(
  Variable = c(
    "Number of participants",
    "Number female (percentage)",
    "Age baseline \u00b1 SD",
    "Follow-up duration (years) \u00b1 SD",
    "Education (years) \u00b1 SD",
    "MMSE baseline\u00b1 SD"
  ),
  `Whole cohort` = c(
    as.character(n_total),
    paste0(n_f_total, " (", pct_f_total, "%)"),
    fmt_ms(summary_df$age_1),
    fmt_ms(summary_df$followup_duration),
    fmt_ms(summary_df$YoE),
    fmt_ms(summary_df$mmse_1)
  ),
  `Non-superagers` = c(
    as.character(n_nsa),
    paste0(n_f_nsa, " (", pct_f_nsa, "%)"),
    fmt_ms(summary_df$age_1[summary_df$group         == "Non-superager"]),
    fmt_ms(summary_df$followup_duration[summary_df$group == "Non-superager"]),
    fmt_ms(summary_df$YoE[summary_df$group           == "Non-superager"]),
    fmt_ms(summary_df$mmse_1[summary_df$group        == "Non-superager"])
  ),
  `Superagers` = c(
    as.character(n_sa),
    paste0(n_f_sa, " (", pct_f_sa, "%)"),
    fmt_ms(summary_df$age_1[summary_df$group         == "Superager"]),
    fmt_ms(summary_df$followup_duration[summary_df$group == "Superager"]),
    fmt_ms(summary_df$YoE[summary_df$group           == "Superager"]),
    fmt_ms(summary_df$mmse_1[summary_df$group        == "Superager"])
  ),
  `p-value` = c(
    "--",
    fmt_p_star(sex_p),
    fmt_p_star(age_baseline_p),
    fmt_p_star(followup_duration_p),
    fmt_p_star(yoe_p),
    fmt_p_star(mmse_p)
  )
)

flextable(tbl2) |>
  bold(part = "header") |>
  autofit()
```

*Abbreviations: SD standard deviation, MMSE Mini Mental State Examination*

**Structure-function coupling, superager status, and episodic memory:** We next used repeated-measures linear mixed-effects models to examine whether SFC varied longitudinally by superager status at the global and network levels. Superagers consistently showed higher SFC than non-superagers globally (β=`r fmt_b(sfc_superager_weighted)`, 95% CI `r fmt_ci_text(sfc_superager_weighted)`, p`r fmt_p(sfc_superager_weighted$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_weighted_mean")`, marginal r²=`r fmt_r2m(sfc_superager_weighted)`, conditional r²=`r fmt_r2c(sfc_superager_weighted)`), in heteromodal regions (β=`r fmt_b(sfc_superager_hmod)`, 95% CI `r fmt_ci_text(sfc_superager_hmod)`, p`r fmt_p(sfc_superager_hmod$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_hmod")`, marginal r²=`r fmt_r2m(sfc_superager_hmod)`, conditional r²=`r fmt_r2c(sfc_superager_hmod)`), and in memory-relevant networks including the DMN (β=`r fmt_b(sfc_superager_dmn)`, 95% CI `r fmt_ci_text(sfc_superager_dmn)`, p`r fmt_p(sfc_superager_dmn$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_dmn")`, marginal r²=`r fmt_r2m(sfc_superager_dmn)`, conditional r²=`r fmt_r2c(sfc_superager_dmn)`) and ECN (β=`r fmt_b(sfc_superager_control)`, 95% CI `r fmt_ci_text(sfc_superager_control)`, p`r fmt_p(sfc_superager_control$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_control")`, marginal r²=`r fmt_r2m(sfc_superager_control)`, conditional r²=`r fmt_r2c(sfc_superager_control)`), with a trend in the SN (β=`r fmt_b(sfc_superager_salience)`, 95% CI `r fmt_ci_text(sfc_superager_salience)`, p`r fmt_p(sfc_superager_salience$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_salience")`, marginal r²=`r fmt_r2m(sfc_superager_salience)`, conditional r²=`r fmt_r2c(sfc_superager_salience)`); however, none of these results survived correction for multiple comparisons. By contrast, SFC in sensory regions did not differ by superager status (p`r fmt_p(sfc_superager_sensory$p_value)`, pFDR`r get_fdr(sfc_fdr_results, "sfc_sensory")`). Figure 3A presents the difference in global SFC between superagers and non-superagers averaged across baseline and follow-up, with warmer colors indicating higher SFC in superagers, and Figure 3B summarizes these results.

**Figure 3** Structure-function coupling, episodic memory, and superagers status analyses  

Figure 3. Structure-function coupling differs by superager status and is associated with episodic memory. (A) Cortical surface and subcortical plot of the unthresholded mean difference in global structure-function coupling between superagers and non-superagers, averaged across both timepoints. Warm colors indicate higher coupling in superagers; cool colors indicate lower coupling. (B, C) Linear mixed-effects model coefficients and 95% confidence intervals for regional structure-function coupling as a function of superager status (B) and episodic memory (C). Filled circles: p < .05 (uncorrected), open circles: non-significant, * pFDR < .05. Abbreviations: FDR false discovery rate, SFC structure-function coupling, DMN default mode network, ECN executive control network, SN salience network, CI confidence interval. 

We next examined the relationship between global and network level SFC and episodic memory across the cohort longitudinally with linear mixed-effects models. Higher SFC was associated with better episodic memory globally (β=`r fmt_b(ravlt_sfc_weighted)`, 95% CI `r fmt_ci_text(ravlt_sfc_weighted)`, p`r fmt_p(ravlt_sfc_weighted$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_weighted_mean")`, marginal r²=`r fmt_r2m(ravlt_sfc_weighted)`, conditional r²=`r fmt_r2c(ravlt_sfc_weighted)`), in heteromodal regions (β=`r fmt_b(ravlt_sfc_hmod)`, 95% CI `r fmt_ci_text(ravlt_sfc_hmod)`, p`r fmt_p(ravlt_sfc_hmod$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_hmod")`, marginal r²=`r fmt_r2m(ravlt_sfc_hmod)`, conditional r²=`r fmt_r2c(ravlt_sfc_hmod)`), and in the DMN (β=`r fmt_b(ravlt_sfc_dmn)`, 95% CI `r fmt_ci_text(ravlt_sfc_dmn)`, p`r fmt_p(ravlt_sfc_dmn$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_dmn")`, marginal r²=`r fmt_r2m(ravlt_sfc_dmn)`, conditional r²=`r fmt_r2c(ravlt_sfc_dmn)`). Associations in the ECN (β=`r fmt_b(ravlt_sfc_control)`, 95% CI `r fmt_ci_text(ravlt_sfc_control)`, p`r fmt_p(ravlt_sfc_control$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_control")`) and SN (β=`r fmt_b(ravlt_sfc_salience)`, 95% CI `r fmt_ci_text(ravlt_sfc_salience)`, p`r fmt_p(ravlt_sfc_salience$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_salience")`) showed trends but did not reach significance. SFC in sensory regions was not associated with episodic memory (p`r fmt_p(ravlt_sfc_sensory$p_value)`, pFDR`r get_fdr(ravlt_fdr_results, "ravlt_sfc_sensory")`). Figure 3C displays these results.

We next examined whether global and network level SFC change – modeled as an interaction with time in linear mixed-effects models – differed by superager status or was associated with episodic memory. A trend emerged in the SN for both superager status (β=`r fmt_b(sfc_superager_salience_slope)`, 95% CI `r fmt_ci_text(sfc_superager_salience_slope)`, p`r fmt_p(sfc_superager_salience_slope$p_value)`, pFDR`r get_fdr(sfc_slope_fdr_results, "sfc_salience_slope")`, marginal r²=`r fmt_r2m(sfc_superager_salience_slope)`, conditional r²=`r fmt_r2c(sfc_superager_salience_slope)`) and episodic memory (β=`r fmt_b(ravlt_sfc_salience_slope)`, 95% CI `r fmt_ci_text(ravlt_sfc_salience_slope)`, p`r fmt_p(ravlt_sfc_salience_slope$p_value)`, pFDR`r get_fdr(ravlt_slope_fdr_results, "ravlt_sfc_salience_slope")`, marginal r²=`r fmt_r2m(ravlt_sfc_salience_slope)`, conditional r²=`r fmt_r2c(ravlt_sfc_salience_slope)`). However, we found no evidence that superager status was associated with SFC change in any other network (ps ≥ `r fmt_round2(min_sfc_slope_p)`, pFDRs ≥ `r fmt_round2(min_sfc_slope_p_fdr)`), or that SFC change was associated with episodic memory (ps ≥ `r fmt_round2(min_ravlt_slope_p)`, pFDRs ≥ `r fmt_round2(min_ravlt_slope_p_fdr)`). Table 2 summarizes all structure-function coupling, superager status, and episodic memory results.

**Table 2** Associations between structure-function coupling, superager status, and episodic memory

```{r table3}
# Helpers to build data rows for LME models
lme_row <- function(region, model, fdr_df, fdr_key) {
  list(
    Region                 = region,
    `Coefficient (95% CI)` = fmt_coef_ci(model),
    `p-value`              = fmt_p_star(as.numeric(model$p_value)[1]),
    pFDR                   = fmt_p_star(get_fdr_raw(fdr_df, fdr_key)),
    `Marginal r²`     = fmt_r2m(model),
    `Conditional r²`  = fmt_r2c(model)
  )
}

# Section header row (spans all columns via merge_at below)
hdr_row <- function(label) {
  list(Region = label, `Coefficient (95% CI)` = "",
       `p-value` = "", pFDR = "",
       `Marginal r²` = "", `Conditional r²` = "")
}

# Section type row (model description label, styled differently from hdr_row)
type_row <- function(label) {
  list(Region = label, `Coefficient (95% CI)` = "",
       `p-value` = "", pFDR = "",
       `Marginal r²` = "", `Conditional r²` = "")
}

tbl3 <- bind_rows(
  # ── Superager status (LME) ──────────────────────────────────────────────
  type_row("Repeated measures linear mixed-effects analyses"),
  hdr_row("Superager status"),
  lme_row("Global SFC",      sfc_superager_weighted, sfc_fdr_results, "sfc_weighted_mean"),
  lme_row("Heteromodal SFC", sfc_superager_hmod,     sfc_fdr_results, "sfc_hmod"),
  lme_row("DMN SFC",         sfc_superager_dmn,      sfc_fdr_results, "sfc_dmn"),
  lme_row("ECN SFC",         sfc_superager_control,  sfc_fdr_results, "sfc_control"),
  lme_row("SN SFC",          sfc_superager_salience, sfc_fdr_results, "sfc_salience"),
  lme_row("Sensory SFC",     sfc_superager_sensory,  sfc_fdr_results, "sfc_sensory"),

  # ── Superager status — change (LME interaction) ─────────────────────────
  type_row("Repeated measures linear mixed-effects analyses — change"),
  hdr_row("Superager status"),
  lme_row("Global SFC × time",      sfc_superager_weighted_slope, sfc_slope_fdr_results, "sfc_weighted_mean_slope"),
  lme_row("Heteromodal SFC × time", sfc_superager_hmod_slope,     sfc_slope_fdr_results, "sfc_hmod_slope"),
  lme_row("DMN SFC × time",         sfc_superager_dmn_slope,      sfc_slope_fdr_results, "sfc_dmn_slope"),
  lme_row("ECN SFC × time",         sfc_superager_control_slope,  sfc_slope_fdr_results, "sfc_control_slope"),
  lme_row("SN SFC × time",          sfc_superager_salience_slope, sfc_slope_fdr_results, "sfc_salience_slope"),
  lme_row("Sensory SFC × time",     sfc_superager_sensory_slope,  sfc_slope_fdr_results, "sfc_sensory_slope"),

  # ── Episodic memory (LME) ───────────────────────────────────────────────
  type_row("Repeated measures linear mixed-effects analyses"),
  hdr_row("Episodic memory"),
  lme_row("Global SFC",      ravlt_sfc_weighted, ravlt_fdr_results, "ravlt_sfc_weighted_mean"),
  lme_row("Heteromodal SFC", ravlt_sfc_hmod,     ravlt_fdr_results, "ravlt_sfc_hmod"),
  lme_row("DMN SFC",         ravlt_sfc_dmn,      ravlt_fdr_results, "ravlt_sfc_dmn"),
  lme_row("ECN SFC",         ravlt_sfc_control,  ravlt_fdr_results, "ravlt_sfc_control"),
  lme_row("SN SFC",          ravlt_sfc_salience, ravlt_fdr_results, "ravlt_sfc_salience"),
  lme_row("Sensory SFC",     ravlt_sfc_sensory,  ravlt_fdr_results, "ravlt_sfc_sensory"),

  # ── Episodic memory — change (LME interaction) ──────────────────────────
  type_row("Repeated measures linear mixed-effects analyses — change"),
  hdr_row("Episodic memory"),
  lme_row("Global SFC × time",      ravlt_sfc_weighted_slope, ravlt_slope_fdr_results, "ravlt_sfc_weighted_mean_slope"),
  lme_row("Heteromodal SFC × time", ravlt_sfc_hmod_slope,     ravlt_slope_fdr_results, "ravlt_sfc_hmod_slope"),
  lme_row("DMN SFC × time",         ravlt_sfc_dmn_slope,      ravlt_slope_fdr_results, "ravlt_sfc_dmn_slope"),
  lme_row("ECN SFC × time",         ravlt_sfc_control_slope,  ravlt_slope_fdr_results, "ravlt_sfc_control_slope"),
  lme_row("SN SFC × time",          ravlt_sfc_salience_slope, ravlt_slope_fdr_results, "ravlt_sfc_salience_slope"),
  lme_row("Sensory SFC × time",     ravlt_sfc_sensory_slope,  ravlt_slope_fdr_results, "ravlt_sfc_sensory_slope")
)

# Identify type rows (model description) and section header rows
type_labels <- c(
  "Repeated measures linear mixed-effects analyses",
  "Repeated measures linear mixed-effects analyses — change"
)
type_idx <- which(tbl3$Region %in% type_labels)
hdr_idx  <- setdiff(which(tbl3$`Coefficient (95% CI)` == ""), type_idx)

ft <- flextable(tbl3) |>
  bold(part = "header") |>
  bold(i = hdr_idx) |>
  bg(i = hdr_idx, bg = "#F0F0F0") |>
  italic(i = type_idx) |>
  bg(i = type_idx, bg = "#E8EEF4") |>
  autofit()

# Merge both type and header rows across all columns so labels span the full width
for (r in c(type_idx, hdr_idx)) {
  ft <- merge_at(ft, i = r, j = 1:6)
}

ft
```

*Abbreviations: CI confidence interval, FDR False Discovery Rate, SFC structure-function coupling, DMN default mode network, ECN executive control network, SN salience network*
